Demers Cartogram with Rivers

### 2023-05-11 paper/section/appendix.tex
draft: VMV 10 page

\section{Pre-processing Shapefiles}
\label{app:pre-processing}
Shapefiles from different sources are likely to be incompatible. In our case, the NHS CCG shapefile is incompatible with the river shapefiles. The major reason for the incompatibility is the coordinate reference system (CRS). The CRS of the CCG shapefile is EPSG:27700 (OSGB36 - British National Grid). The CRS of the river shapefiles is EPSG:4326 (WGS84 - World Geodetic System). Here we provide some pre-processing steps using QGIS (version: 3.26.0-Buenos Aires) \cite{qgisWelcome} to handle the incompatibility issue and reduce shapefile size to improve performance.

@@

{
\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{figure/qgis/import_rivers.png}
    \caption{QGIS interface, with River Trent, River Great Ouse, and River Thames (from top to bottom) imported.}
    \label{fig:import_rivers}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{figure/qgis/import_ccgs.png}
    \caption{QGIS interface, with all NHS CCGs imported.}
    \label{fig:import_ccgs}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{figure/qgis/unify_crs.png}
    \caption{QGIS interface, showing the unified CRS (OSGB36) for both layers.}
    \label{fig:unify_crs}
\end{figure}
}

@@

{
\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{figure/qgis/export_rivers.png}
    \caption{QGIS interface, exporting all rivers using the OSGB36 CRS in GeoJSON.}
    \label{fig:export_rivers}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{figure/qgis/export_ccgs.png}
    \caption{QGIS interface, exporting all NHS CCGs using the OSGB36 CRS in GeoJSON.}
    \label{fig:export_ccgs}
\end{figure}

@@

{
\begin{figure}[h!]
    \centering
    \includegraphics[width=\columnwidth]{figure/qgis/export_topojson.png}
    \caption{Mapshaper interface, merging all rivers with NHS CCGs into one layer, and export the merged layer in TopoJSON.}
    \label{fig:export_topojson}
\end{figure}
}

@@

    {
        \renewcommand{\arraystretch}{1.5}
        \begin{table}[!tb]
            \centering
            \resizebox{\columnwidth}{!}{
                \begin{tabulary}{\columnwidth}{|*{4}{l|}}
                    \hhline{~|*{3}{-}}
                    \multicolumn{1}{c|}{\textbf{Shapefile}} &
                    \cellcolor{Mycolor2}\textbf{Original} &
                    \cellcolor{Mycolor2}\textbf{GeoJSON} &
                    \cellcolor{Mycolor2}\textbf{TopoJSON} \\
                    \hline
                    Rivers & 2.0 MB (GeoJSON) & 1.4 MB & -  \\
                    \hline
                    NHS CCGs & 46.6 MB (.shp, Esri vector shapefile) & 140.2 MB & -  \\
                    \hline
                    Merged & - & - & 16.3 MB \\
                    \hline
                \end{tabulary}
            }
            \caption{The file size is reduced by 88.5\% from the original size.}
            \label{table:pre-processing_result}
        \end{table}
    }

@@

@@

\begin{algorithm}[h!]
    \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
    \textbf{Input:} \\
    $ \nodeList \gets $ a list of $ \node $ representing CCGs \\

    \textbf{Local variables:} \\
    $ \riverList \gets $ a list of $ \river $ representing river features \\
    $ \node \gets $ a node is an object with the following properties: \\
    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
    $ j \gets $ the number of nodes intersecting $ \river $  \\

    \begin{algorithmic}[1]
        \Procedure{TranslateRiver}{$ \nodeList $}
        \ForEach {$ \river \in \riverList $}
        \State $ j \gets 0 $
        \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
        \ForEach {$ \node \in \nodeList $}

        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
        \If{ \Call{TestIntersection}{$ \nodeFNORLine $, $ \river $} $ = True $}
        \State $ \riverVector \gets \riverVector + \nodeVectorCT$
        \State $ j ++ $
        \EndIf

@@

\section{Procedure: TestIntersection}
% \begin{noindent}
    \begin{algorithm}[h!]
        \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}
        \textbf{Input:} \\
        $ \nodeLineNV \gets $ the node's translation path \\
        $ \river \gets $ a river feature \\

@@

    \begin{algorithm}[h!]
        \caption{Procedure to derive a point based an edge and a distance.}\label{alg:derive corridor point}
        \textbf{Input:} \\
        $ \Edge \gets $ the edge used to derive the new point \\
        $ \Distance \gets $ the distance between $ \PointP $ and $ \EdgeStart $ \\

@@

\begin{algorithm}[h!]
    \caption{Procedure to derive an edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $.}\label{alg:derive corridor edge}

>>>>

\let\origappendixsection=\section
\renewcommand\section[1]{\origappendixsection{\textbf{#1}}}

\section{Obtaining Shapefile}
\label{obtaining-shapefile}
\begin{lstlisting}[caption={The query that downloads the shapefile of River Thames from OpenStreetMap via the Overpass Turbo API.}, label={overpass},captionpos=b]
    relation(2263653);>>;
    // River Great Ouse: 2798097
    // River Trent: 2863468
    out skel;
\end{lstlisting}
\section{Pre-processing Shapefiles}
\label{app:pre-processing}
Shapefiles from different sources are likely to be incompatible. In our case, the NHS CCG shapefile is incompatible with the river shapefiles. The major reason for the incompatibility is the coordinate reference system (CRS). The CRS of the CCG shapefile is EPSG:27700 (OSGB36 - British National Grid). The CRS of the river shapefiles is EPSG:4326 (WGS84 - World Geodetic System). Here we provide some pre-processing steps using QGIS (version: 3.26.0-Buenos Aires) \cite{qgisWelcome} to handle the incompatibility and reduce shapefile size to improve performance.

@@

{
\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{figure/qgis/import_rivers.png}
    \caption{QGIS interface, with River Trent, River Great Ouse, and River Thames (from top to bottom) imported.}
    \label{fig:import_rivers}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{figure/qgis/import_ccgs.png}
    \caption{QGIS interface, with all NHS CCGs imported.}
    \label{fig:import_ccgs}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{figure/qgis/unify_crs.png}
    \caption{QGIS interface, showing the unified CRS (OSGB36) for both layers.}
    \label{fig:unify_crs}
\end{figure}
}

@@

{
\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{figure/qgis/export_rivers.png}
    \caption{QGIS interface, exporting all rivers using the OSGB36 CRS in GeoJSON.}
    \label{fig:export_rivers}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{figure/qgis/export_ccgs.png}
    \caption{QGIS interface, exporting all NHS CCGs using the OSGB36 CRS in GeoJSON.}
    \label{fig:export_ccgs}
\end{figure}

@@

{
\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{figure/qgis/export_topojson.png}
    \caption{Mapshaper interface, merging all rivers with NHS CCGs into one layer, and export the merged layer in TopoJSON.}
    \label{fig:export_topojson}
\end{figure}
}

@@

    {
        \renewcommand{\arraystretch}{1.5}
        \begin{table}[H]
            \centering
            \resizebox{\columnwidth}{!}{
                \begin{tabulary}{\columnwidth}{|*{4}{l|}}
                    \hhline{~|*{3}{-}}
                    \multicolumn{1}{c|}{\textbf{Shapefile}} &
                    \cellcolor{Mycolor2}\textbf{Original} &
                    \cellcolor{Mycolor2}\textbf{GeoJSON} &
                    \cellcolor{Mycolor2}\textbf{TopoJSON} \\
                    \hline
                    Rivers & 2.0 MB (GeoJSON) & 1.4 MB & -  \\
                    \hline
                    NHS CCGs & 46.6 MB (.shp, Esri vector shapefile) & 140.2 MB & -  \\
                    \hline
                    Merged & - & - & 16.3 MB \\
                    \hline
                \end{tabulary}
            }
            \caption{The file size is reduced by 88.5\% from the original size.}
            \label{table:pre-processing_result}
        \end{table}
    }

@@

\section{Procedure: TranslateNode}

\begin{algorithm}[H]
    \caption{Procedure to translate node positions.}\label{alg:TranslateNode}

    \textbf{Input:} \\
    $ \nodeList \gets $ a list of $ \node $ representing regions \\

    \textbf{Output:} \\
    $ \nodeTopologicalError \gets $ the number of river crossing nodes in the input \\

    \textbf{Global variables:} \\
    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\

    \textbf{Local variables:} \\
    $ \node \gets $ a node is an object with the following properties: \\
    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river  \\
    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\

    \begin{algorithmic}[1]
        \Procedure{TranslateNode}{$ \nodeList $}
        \State $ \nodeTopologicalError \gets 0 $
        \ForEach {$ \node \in \nodeList $}

        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}

        \State $ \node(x,y) \gets \nodeFNOR(x,y) $

        \If{ \Call{TestIntersection}{$ \nodeFNORLine $} $ = True $}
        \State $ \nodeCross ++ $
        \State $ \nodeTopologicalError ++ $

        \If{$ \nodeCross < \stalemateMax $}
        \State \hskip-2em \Comment{Translate back to previous position}
        \State $ \node(x,y) \gets \nodePrevious(x,y) $

        \Else

        \State \Call{ProcessStalemate}{$ \node, \nodeFNOR $}
        \State $ \nodeCross \gets 0 $ \Comment{Reset counter}

        \EndFor
        \State \Return{$ \nodeTopologicalError $}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\section{Procedure: UpdateLayout}

\begin{algorithm}[H]
    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \Cref{sec:{Demers Cartogram with Rivers}}.}\label{alg:UpdateLayout}
    \textbf{Global variables:} \\
    $ \nodeList \gets $ a list of $ \node $ representing regions with the following properties: \\
    \-\hspace{1em}  $\nodeListCross \gets $ the number of $ \node $ in $ \nodeList $ that crosses a river \\

    $ \nodeSize \gets $ the initial size of all nodes \\
    $ \nodeCartographicError \gets $ the average cartographic error of all nodes \\
    $ \nodeCartographicErrorMax \gets $ the maximum cartographic error of all nodes \\
    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\

    \textbf{Local variables:} \\
    $ \node \gets $ a node is an object with the following properties: \\
    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river \\
    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\

    \begin{algorithmic}[1]
        \Procedure{UpdateLayout}{}

        \State $ \nodeSize \gets 1 $
        \While{$ \nodeCartographicError < \nodeCartographicErrorMax $}
        \State $ \nodeListCross \gets 1 $ \Comment{Trigger the while loop}
        \While{$ \nodeListCross > 0 $}

        \State $ \nodeListCross \gets 0 $
        \State $ \nodeList \gets $ RemoveOverlap ($ \nodeList $)
        \State \Call{TranslateRiver}{$ \nodeList $}
        \State $\nodeListCross \gets $ \Call{TranslateNode}{$ \nodeList $}

        \EndWhile

        \State $ \nodeSize ++ $

        \EndWhile

@@

\begin{algorithm}[H]
    \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
    \textbf{Input:} \\
    $ \nodeList \gets $ a list of $ \node $ representing regions \\

    \textbf{Local variables:} \\
    $ \riverList \gets $ a list of $ \river $ representing river features \\
    $ \node \gets $ a node is an object with the following properties: \\
    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
    $ \nodeTopologicalError \gets $ the number of nodes intersecting $ \river $  \\

    \begin{algorithmic}[1]
        \Procedure{TranslateRiver}{$ \nodeList $}
        \ForEach {$ \river \in \riverList $}
        \State $ \nodeTopologicalError \gets 0 $
        \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
        \ForEach {$ \node \in \nodeList $}

        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
        \If{ \Call{TestIntersection}{$ \nodeFNORLine $, $ \river $} $ = True $}
        \State $ \riverVector \gets \riverVector + \nodeVectorCT$
        \State $ \nodeTopologicalError ++ $
        \EndIf

@@

\section{Procedure: ProcessStalemate}

    \begin{algorithm}[H]
        \caption{Procedure to derive a corridor to resolve stalemate. We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards (See \Cref{fig:stalemate}).}\label{alg:derive corridor}

        \textbf{Input:} \\
        $ \node \gets $ the node used to derive the corridor \\

        \textbf{Global variables:} \\
        $ \CorridorLength \gets $ the length of a corridor \\
        $ \CorridorWidth \gets $ the width of a corridor \\

        \textbf{Local variables:} \\
        $ \Corridor \gets $ the corridor \\
        $ \PointP \gets $ the point extending $ \nodeVectorTC $ such that $ \nodeLineWidthNtP = \CorridorLength $\\
        $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \nodeLineNtNc $ \\
        $ corridor \gets $ a rectangle formed by $ \EdgeParallelA $ and $ \EdgeParallelB $ \\

        \begin{algorithmic}[1]
            \Procedure{ProcessStalemate}{$ \node $}
                \State $ \node(x,y) \gets \nodePrevious(x,y) $

                \State $ \PointP \gets $ \Call{DerivePoint}{$ \nodeVectorTC $, $ \CorridorLength $}

                \State $ \EdgeParallelA \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ \frac{\CorridorWidth}{2} $}

                \State $ \EdgeParallelB \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ -\frac{\CorridorWidth}{2} $}

                \State $ \Corridor \gets
                    \begin{bmatrix}
                        \EdgeParallelA.start &
                        \EdgeParallelA.end \\

                        \EdgeParallelB.start &
                        \EdgeParallelB.end \\
                    \end{bmatrix} $

                \ForEach{$ \nodeInCorridor $ inside $ \Corridor $}

                    \State $ \Vector{\nodeInCorridor ~\nodeInCorridorT} = \nodeVectorTC $

                    \State $ \nodeInCorridor(x,y) \gets \nodeInCorridorT(x,y) $

                \EndFor
            \EndProcedure
        \end{algorithmic}
    \end{algorithm}

    %\end{noindent}
\section{Procedure: TestIntersection}
% \begin{noindent}
    \begin{algorithm}[H]
        \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}
        \textbf{Input:} \\
        $ \nodeLineNV \gets $ the node's translation path \\
        $ \river \gets $ a river feature \\

@@

    \begin{algorithm}[H]
        \caption{Procedure to derive a point based an edge and a distance.}\label{alg:derive corridor point}
        \textbf{Input:} \\
        $ \Edge \gets $ the edge used to derive the new point \\
        $ \Distance \gets $ the distance between $ \PointP $ and $ \EdgeStart $ \\

@@

\begin{algorithm}[H]
    \caption{Procedure to derive an edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $.}\label{alg:derive corridor edge}

### 2023-05-11 paper/section/data.tex
draft: review
draft: VMV 10 page

 {
  \begin{figure}[b!]
      \centering
      \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
      \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography Portalx \cite{opengeographyportalxOpen} with EPSG:4326 (WGS84 - World Geodetic System) as the Coordinate Reference System (CRS).}
      \label{fig:ccg}
  \end{figure}
 }

@@

\bobgraph{Choropleth Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}. The number of CCGs changes over time due to NHS re-organization. The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}. We decided to use the CCG shapefile from 2020 at the time of writing (See \Cref{fig:ccg}), due to the absence of public EHR data published based on the latest CCG re-organizations that took place in 2021 and 2022.

\bobgraph{River Shapefiles:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent, and River Great Ouse in England. These rivers were chosen as they are well-known rivers and pass through regions with dense populations, and provide informative geographical and topological cues. While including smaller rivers is technically feasible it may not increase the legibility of the cartogram.

We first obtain a relation ID by searching for a river, e.g. River Thames, on OpenStreetMap. The relation ID is used to construct a query (See Listing~\ref{overpass}) which enables the user to download the entire river shapefile using Overpass Turbo \cite{overpassturboOverpass}.

\begin{lstlisting}[float=tp,caption={The query that downloads the shapefile of River Thames from OpenStreetMap via the Overpass Turbo API.}, label={overpass},captionpos=b]
    relation(2263653);>>;
    // River Great Ouse: 2798097
    // River Trent: 2863468
    out skel;
\end{lstlisting}

>>>>

%  {
%   \begin{figure}[b!]
%       \centering
%       \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
%       \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography Portalx \cite{opengeographyportalxOpen} with EPSG:4326 (WGS84 - World Geodetic System) as the Coordinate Reference System (CRS).}
%       \label{fig:ccg}
%   \end{figure}
%  }

@@

\bobgraph{Choropleth Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}. The number of CCGs changes over time due to NHS re-organization. The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}. We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG re-organizations that took place in 2021 and 2022.

\bobgraph{River Shapefiles:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent, and River Great Ouse in England. These rivers were chosen as they are well-known rivers and pass through regions with dense populations, and provide informative geographical and topological cues. While including smaller rivers is technically feasible it may not increase the legibility of the cartogram. This is an open question for future work.

We first obtain a relation ID by searching for a river, e.g. River Thames, on OpenStreetMap. The relation ID is used to construct a query (See \cref{obtaining-shapefile}) which enables the user to download the entire river shapefile using Overpass Turbo \cite{overpassturboOverpass}.

### 2023-05-11 paper/section/main.tex
draft: VMV 10 page

\section{Demers Cartogram with Rivers}
 {
  \begin{figure*}[tb!]
      \centering
      \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figure/flowchart.png}
      \caption{An overview of our hybrid layout algorithm incorporating rivers. See also \Cref{alg:UpdateLayout} in \Cref{sec:{Demers Cartogram with Rivers}} for more detail. See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate. For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
      \label{fig:flowchart}
  \end{figure*}
 }

@@

\begin{algorithm}[tb!]
    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \Cref{sec:{Demers Cartogram with Rivers}}.}\label{alg:UpdateLayout}
    \textbf{Global variables:} \\
    $ \nodeList \gets $ a list of $ \node $ representing regions with the following properties: \\
    \-\hspace{1em}  $\nodeListCross \gets $ the number of $ \node $ in $ \nodeList $ that crosses a river \\

    $ \nodeSize \gets $ the initial size of all nodes \\
    $ \nodeCartographicError \gets $ the average cartographic error of all nodes \\
    $ \nodeCartographicErrorMax \gets $ the maximum cartographic error of all nodes \\
    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\

    \textbf{Local variables:} \\
    $ \node \gets $ a node is an object with the following properties: \\
    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river \\
    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\

    \begin{algorithmic}[1]
        \Procedure{UpdateLayout}{}

        \State $ \nodeSize \gets 1 $
        \While{$ \nodeCartographicError < \nodeCartographicErrorMax $}
        \State $ \nodeListCross \gets 1 $ \Comment{Trigger the while loop}
        \While{$ \nodeListCross > 0 $}

        \State $ \nodeListCross \gets 0 $
        \State $ \nodeList \gets $ RemoveOverlap ($ \nodeList $)
        \State \Call{TranslateRiver}{$ \nodeList $}
        \State $\nodeListCross \gets $ \Call{TranslateNode}{$ \nodeList $}

        \EndWhile

        \State $ \nodeSize ++ $

        \EndWhile

        \EndProcedure
    \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[tb!]
    \caption{Procedure to translate node positions.}\label{alg:TranslateNode}

    \textbf{Input:} \\
    $ \nodeList \gets $ a list of $ \node $ representing regions \\

    \textbf{Output:} \\
    $ \nodeTopologicalError \gets $ the number of river crossing nodes in the input \\

    \textbf{Global variables:} \\
    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\

    \textbf{Local variables:} \\
    $ \node \gets $ a node is an object with the following properties: \\
    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river  \\
    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\

    \begin{algorithmic}[1]
        \Procedure{TranslateNode}{$ \nodeList $}
        \State $ \nodeTopologicalError \gets 0 $
        \ForEach {$ \node \in \nodeList $}

        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}

        \State $ \node(x,y) \gets \nodeFNOR(x,y) $

        \If{ \Call{TestIntersection}{$ \nodeFNORLine $} $ = True $}
        \State $ \nodeCross ++ $
        \State $ \nodeTopologicalError ++ $

        \If{$ \nodeCross < \stalemateMax $}
        \State \hskip-2em \Comment{Translate back to previous position}
        \State $ \node(x,y) \gets \nodePrevious(x,y) $

        \Else

        \State \Call{ProcessStalemate}{$ \node, \nodeFNOR $}
        \State $ \nodeCross \gets 0 $ \Comment{Reset counter}

        \EndIf

        \EndIf

        \EndIf

        \EndFor
        \State \Return{$ \nodeTopologicalError $}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[tbh!]
    \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
    \textbf{Input:} \\
    $ \nodeList \gets $ a list of $ \node $ representing regions \\

    \textbf{Local variables:} \\
    $ \riverList \gets $ a list of $ \river $ representing river features \\
    $ \node \gets $ a node is an object with the following properties: \\
    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
    $ \nodeTopologicalError \gets $ the number of nodes intersecting $ \river $  \\

    \begin{algorithmic}[1]
        \Procedure{TranslateRiver}{$ \nodeList $}
        \ForEach {$ \river \in \riverList $}
        \State $ \nodeTopologicalError \gets 0 $
        \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
        \ForEach {$ \node \in \nodeList $}

        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
        \If{ \Call{TestIntersection}{$ \nodeFNORLine $, $ \river $} $ = True $}
        \State $ \riverVector \gets \riverVector + \nodeVectorCT$
        \State $ \nodeTopologicalError ++ $
        \EndIf

        \EndIf

        \EndFor

        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{\nodeTopologicalError} $
        \EndFor

        \EndProcedure
    \end{algorithmic}
\end{algorithm}

@@

{
    \begin{figure}[tb!]
        \centering
        \includegraphics[width=\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
        \caption{A flowchart illustration of stalemate processing. See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
        \label{fig:flowchart-stalemate}
    \end{figure}
   }

@@

% \begin{noindent}

\begin{algorithm}[tb!]
    \caption{Procedure to derive a corridor to translate enclosed nodes. We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards (See \Cref{fig:stalemate}).}\label{alg:derive corridor}

    \textbf{Input:} \\
    $ \node \gets $ the node used to derive the corridor \\

    \textbf{Global variables:} \\
    $ \CorridorLength \gets $ the length of a corridor \\
    $ \CorridorWidth \gets $ the width of a corridor \\

    \textbf{Local variables:} \\
    $ \Corridor \gets $ the corridor \\
    $ \PointP \gets $ the point extending $ \nodeVectorTC $ such that $ \nodeLineWidthNtP = \CorridorLength $\\
    $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \nodeLineNtNc $ \\
    $ corridor \gets $ a rectangle formed by $ \EdgeParallelA $ and $ \EdgeParallelB $ \\

    \begin{algorithmic}[1]
        \Procedure{ProcessStalemate}{$ \node $}
            \State $ \node(x,y) \gets \nodePrevious(x,y) $

            \State $ \PointP \gets $ \Call{DerivePoint}{$ \nodeVectorTC $, $ \CorridorLength $}

            \State $ \EdgeParallelA \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ \frac{\CorridorWidth}{2} $}

            \State $ \EdgeParallelB \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ -\frac{\CorridorWidth}{2} $}

            \State $ \Corridor \gets
                \begin{bmatrix}
                    \EdgeParallelA.start &
                    \EdgeParallelA.end \\

                    \EdgeParallelB.start &
                    \EdgeParallelB.end \\
                \end{bmatrix} $

            \ForEach{$ \nodeInCorridor $ inside $ \Corridor $}

                \State $ \Vector{\nodeInCorridor ~\nodeInCorridorT} = \nodeVectorTC $

                \State $ \nodeInCorridor(x,y) \gets \nodeInCorridorT(x,y) $

            \EndFor
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

%\end{noindent}

@@

{
\begin{figure*}[tb!]
    \centering
    \includegraphics[width=\textwidth,keepaspectratio]{figure/UI.png}
    \caption{A screenshot of the user interface. User options are provided to adjust the terminating conditions (size and error), color mapping, visibility of nodes and rivers. Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic. See \Cref{subsec:{User Options}} for more on user options. In this figure, $ \nodeCartographicErrorMax $ = 1.875\%, and an $ \nodeTopologicalError $ of 9.63\% is eliminated.}
    \label{fig:overview}
\end{figure*}
}

>>>>

\section{Demers Cartogram with Rivers}
 {
  \begin{figure*}[tb!]
      \centering
      \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figure/flowchart.png}
      \caption{An overview of our hybrid layout algorithm incorporating rivers. See also \Cref{alg:UpdateLayout} in \Cref{sec:{Demers Cartogram with Rivers}} for more detail. % See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
       For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
      \label{fig:flowchart}
  \end{figure*}
 }

@@

@@

@@

@@

% {
%     \begin{figure}[tb!]
%         \centering
%         \includegraphics[width=\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
%         \caption{A flowchart illustration of stalemate processing. See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
%         \label{fig:flowchart-stalemate}
%     \end{figure}
%    }

@@

@@

{
\begin{figure*}[tb!]
    \centering
    \includegraphics[width=0.7\textwidth,keepaspectratio]{figure/UI.png}
    \caption{A screenshot of the user interface. User options are provided to adjust the terminating conditions (size and error), color mapping, visibility of nodes and rivers. Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic. See \Cref{subsec:{User Options}} for more on user options. In this figure, $ \nodeCartographicErrorMax $ = 1.875\%, and an $ \nodeTopologicalError $ of 9.63\% is eliminated.}
    \label{fig:overview}
\end{figure*}
}
